---
title: "HW7"
author: "Zachary Brunell"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: true
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height=8, fig.width= 12, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60), message=FALSE)
```

# [GitHub Link](https://github.com/zbrunell/SDS315/blob/main/HW7.Rmd)
\newpage

# Problem 1

## Part A

```{r Problem 1, echo=FALSE, results = 'asis'}
library(ggplot2)
library(mosaic)
library(dplyr)
library(tidyverse)
library(readr)
library(MatchIt)

arms= read_csv('armfold.csv')
male = sum(arms$Sex == 'Male')
female = sum(arms$Sex == 'Female')

male_folded = sum((arms$Sex == 'Male') & (arms$LonR_fold == 1))
female_folded = sum((arms$Sex == 'Female') & (arms$LonR_fold == 1))
male_folded_pct = round((male_folded / male), 4)
female_folded_pct = round((female_folded / female), 4)
obs_diff = male_folded_pct - female_folded_pct
cat("There are", male, "males and", female, "females.\n\n")
cat("Sample proportion of males who folded their left arm on top:", paste0(male_folded_pct), "\n\n")
cat("Sample proportion of females who folded their left arm on top:", paste0(female_folded_pct), "\n\n")
```

## Part B

```{r Problem 1B, echo=FALSE, results = 'asis'}
cat("Observed difference in proportions (males minus females):", paste0(obs_diff), "\n\n")
```

## Part C (Using R's Built-In Prop.Test)

```{r Problem 1C, echo=FALSE, results = 'show'}
boot_arms = prop.test(LonR_fold~Sex, data= arms)
confint(boot_arms, level = 0.95)
# Hand calculated
p1 = round(male_folded/male, 4)
p2 = round(female_folded/female, 4)
SE_diff = sqrt((p1 * (1-p1)/male) + (p2 * (1 - p2)/female))
z_score = 1.96
obs_diff = p1 - p2
margin_of_error = z_score * SE_diff
lower_bound <- obs_diff - margin_of_error
upper_bound <- obs_diff + margin_of_error
```
This calculates the 95% confidence interval for the difference in proportions between males and females who folded their left hand on top using prop.test to be -0.09 to 0.19.

## Part C (Hand-Written)

Formula for Standard Error: \Large $SE(\hat{p}_1 - \hat{p}_2) = \sqrt{\frac{\hat{p}_1(1 - \hat{p}_1)}{N_1} + \frac{\hat{p}_2(1 - \hat{p}_2)}{N_2}}$ \normalsize

Values plugged into the formula: $P_1$ = `r p1`, $P_2$ = `r p2`, $N_1$ = `r male`, $N_2$ = `r female`

Z-score used for 95% confidence: `r z_score`.
This is because in a normal distribution, 95% of the data lies within around 1.96 standard deviations from the mean.

Standard error: `r round(SE_diff, 4)`

Margin of error: `r round(margin_of_error, 4)`

95% Confidence Interval for the difference in proportions: [`r round(lower_bound, 4)` , `r round(upper_bound, 4)`]

## Part D

If we were to repeat this experiment many times, 95% of the time the true difference in proportions (between males and females) would fall within the interval of [`r round(lower_bound, 4)`, `r round(upper_bound, 4)`].

## Part E

The standard error represents how much the difference in proportions (between males and females) is likely to vary across different random samples drawn from the population.
It measures the expected spread of the difference in proportions if we were to repeatedly sample from the population, as in bootstrapping.

## Part F

A sampling distribution refers to distribution of the difference in sample proportions.
It shows how the statistic would vary across samples.
The proportion of males and females who folded their left arm over their right will vary, and therefore, so will the difference in the proportion of those two.
This is due to the nature of random sampling.
However, the true population proportions of the males and females who folded their left arm over their right stays fixed.
This is the "true" values we would discover if we could somehow repeat this experiment for all men and women.

## Part G

The mathematical theorem which justifies the use of a normal distribution is the Central Limit Theorem (heretofore referred to as the CLT).
The CLT states that for sufficiently large sample sizes (NP \>= 10 and N(1-P) \>= 10), the sampling distribution of the sample mean/proportion is approximately normal, regardless of the shape of the population's distribution.
Since $N_1$($P_1$) = `r male * p1`, which is greater than or equal to 1010 and $N_1$(1-$P_1$) = `r male * (1-p1)`, which is greater than or equal to 10, this theorem holds for the male proportions.
Since $N_2$($P_2$) = `r female * p2`, which is greater than or equal to 10 and $N_2$(1-$P_2$) = `r female * (1-p2)`, which is greater than or equal to 10, this theorem holds for females too.
Thus, it holds for the difference of the two sample proportions.

## Part H

If the interval were [-0.01, 0.30], it would be fair to say that there is no difference in arm folding between sexes.
This is because the null hypothesis is there is no difference between the sexes, and due to 0 being inside this interval, we would fail to reject this hypothesis (due to lack of statisically significant evidence).
Thus, there would not be statistical evidence of any difference.

## Part I

If this experiment were conducted many times with different random samples of university students, it is highly likely that the confidence intervals would differ across samples.
This is because each random sample introduces some variability in the sample difference in proportions, and as a result, the values will change with each new sample.
In any random sample, there is a degree of randomness, and as such, the sample proportions will vary slightly.
This variability is often referred to as Monte Carlo variation.
However, we can be confident that approximately 95% of the confidence intervals would contain the true difference in proportions for the population (all men and women).
\newpage
# Problem 2

## Part A
```{r echo=FALSE}
turnout = read_csv('turnout.csv')
```
```{r echo=FALSE}
gotv_call = prop(voted1998 ~ GOTV_call, data = turnout)

gotv_call_props <- as.data.frame(t(gotv_call))
colnames(gotv_call_props) <- c("0", "1")
```

The proportion of individuals who received a GOTV call and voted in 1998 was: **`r gotv_call_props$'1'`**

The proportion of individuals who did not receive a GOTV call and voted in 1998 was: **`r gotv_call_props$'0'`**

```{r echo=FALSE, show_col_types = FALSE}
boot_gotv = prop.test(voted1998 ~ GOTV_call, data = turnout)
confint(boot_gotv)
```

Using R’s built-in prop.test, the 95% confidence interval for the difference in voting proportions between those who received a GOTV call and those who did not is [0.141, 0.266]. Since this interval does not contain zero, the result is statistically significant, suggesting that receiving a GOTV call is associated with higher voter turnout. However, to determine whether this relationship is due to underlying factors, we must examine potential confounders by comparing group means across relevant variables.

## Part B

```{r Problem 2, echo=FALSE, results = 'show', show_col_types = FALSE}
turnout = read_csv('turnout.csv')
voted_summary = turnout %>%
  group_by(GOTV_call) %>%
  summarize( Age = mean(AGE), 'Voted in 1996' = mean(voted1996),"Major Party" = mean(MAJORPTY))
voted_summary

```

The summary statistics reveal a clear difference in proportions between individuals who received a GOTV call and those who did not. For example, the mean age of those who received a call was approximately 58, compared to only 49 for those who did not. Given that older individuals are generally more likely to vote, age serves as a confounding variable in this context. This pattern holds for other variables as well, such as whether the individual voted in 1996 and whether they are affiliated with a major party. Since these variables are associated with both the treatment and the outcome, they qualify as confounders.

### Age

```{r echo=FALSE, show_col_types = FALSE}
boot_ages = t.test(AGE ~ GOTV_call, data = turnout)
confint(boot_ages)
ggplot(turnout) +
  geom_histogram(aes(x=AGE,y=after_stat(density)),binwidth=2) +
  facet_wrap(~GOTV_call) + labs(x="Ages",y="Frequency (proportion)",title="Age Distribution by GOTV call")
```
Furthermore, the 95% confidence interval for the difference in mean ages between individuals who did not receive a GOTV call and those who did ranges from -11.40 to -6.37. Since this interval does not include zero, the difference is statistically significant, suggesting that age may be a confounding variable in this context. This conclusion is further supported by the visualizations, which clearly illustrate a substantial age gap between the two groups.

\newpage

### Voted in 1996
```{r echo=FALSE, show_col_types = FALSE}
boot_1996 = prop.test(voted1996 ~ GOTV_call, data = turnout)
confint(boot_1996)
```

The 95% confidence interval here is [0.122, 0.241]. Since this interval does not include 0, it is statistically significant, meaning it supports the conclusion that voting in 1996 is a confounding variable.

\newpage

### Major Party Affiliation

```{r echo=FALSE, show_col_types = FALSE}
boot_party = prop.test(MAJORPTY ~ GOTV_call, data = turnout)
confint(boot_party)
```

The 95% confidence interval for the difference in proportions between individuals who were registered with a major party and who did or did not receive a GOTV_call ranges from 0.004 to 0.109. Because this interval does not include zero, the result is statistically significant. This supports the claim of a relationship between receiving a GOTV call and major party affiliation.

\newpage

## Part C

```{r Problem 2b, echo=FALSE, results = 'hide', show_col_types = FALSE}
turnout_matched = matchit(GOTV_call ~ voted1996 + AGE + MAJORPTY,
                                 data = turnout,
                            method = 'optimal',
                           ratio = 5)
matched_data <- match.data(turnout_matched)
```
```{r echo = FALSE, show_col_types = FALSE}
summary_stats = matched_data %>%
  group_by(GOTV_call) %>%
  summarize('Voted in 1996' = mean(voted1996), Age = mean(AGE),'Major Party' = mean(MAJORPTY))
summary_stats
```
The above histogram depicts the similarities in our age variable and distribution after matching. This shows almost zero difference between the two due to our matching, which might mean we no longer have tese confounders. To further prove this, we will again look at means across the two groups.
### Age (Matched)
```{r echo=FALSE, show_col_types = FALSE}
matched_age = t.test(AGE ~ GOTV_call, data = matched_data)
confint(matched_age)

ggplot(matched_data) +
  geom_histogram(aes(x=AGE ,y=after_stat(density)),binwidth=2) +
  facet_wrap(~GOTV_call) + labs(x="Ages",y="Frequency (proportion)",title="Age Distribution by GOTV call (received or did not)")
```

When examining our 95% confidence interval, it is clear that the difference in mean ages for the matched data set falls between -1.74 and 3.71. Because this interval includes 0, the result is not statistically significant—unlike before—and thus supports the conclusion that age is no longer acting as a confounding variable in the matched data.

\newpage

### Voted in 1996 (Matched)

```{r echo=FALSE, show_col_types = FALSE}
matched_voted1996 = prop.test(voted1996 ~ GOTV_call, data = matched_data)
confint(matched_voted1996)
```
The 95% confidence interval for the difference in proportions of those who voted in 1996 and did/did not receive GOTV_call to illustrate this difference ranged from -0.03 to 0.10, meaning it is not statistically significant, further supporting the claim that the treatment (GOVT_call) is not associated with being registered for a major party in our matched data set.

### Major Party (Matched)

```{r echo=FALSE, show_col_types = FALSE}
matched_party = prop.test(MAJORPTY ~ GOTV_call, data = matched_data)
confint(matched_party)
```

The 95% confidence interval for the difference in proportions of individuals registered with a major party between those who did and did not receive a GOTV call ranges from -0.05 to 0.07. Since this interval includes zero, the difference is not statistically significant. This supports the conclusion that, in our matched data set, receiving a GOTV call is not associated with major party registration.

\newpage

## Wrap-Up

```{r echo=FALSE, show_col_types = FALSE}
prop(voted1998 ~ GOTV_call, data= matched_data)
```

Using the matched data set, we observe that individuals who received a GOTV call were more likely to vote compared to those who did not. To support this observation with statistical evidence, we will conduct a prop.test, which uses the principles of repeated random sampling to estimate a confidence interval for the difference in proportions and assess statistical significance.

```{r echo=FALSE, show_col_types = FALSE}
matched_GOTV = prop.test(voted1998 ~ GOTV_call, data = matched_data)
confint(matched_GOTV)
```

Since the 95% confidence interval for the difference in proportions for the voting in 1998 and whether or not someone received a GOTV_call ranges from 0.04 and 0.17 (does not include zero), there is statistically significant evidence that receiving a call increases the rate of voting.
To wrap this up, we can quite confidently state that the GOTV_call had essentially no effect on the likelihood/proportion of voting in the 1998 election. Due to our matching, age, party affiliation proportions, and voting in 1996 are no longer confounders, which allows us to see the true relationship between GOTV and voting in 1998. 
