---
title: "HW7"
author: "Zachary Brunell"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height=8, fig.width= 12, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60), message=FALSE)
```
# [GitHub Link](https://github.com/zbrunell/SDS315/compare/main...HW7?expand=1)

# Problem 1
```{r Problem 1, echo=FALSE, results = 'asis'}
library(ggplot2)
library(mosaic)
library(dplyr)
library(tidyverse)
library(readr)
library(MatchIt)

arms= read_csv('armfold.csv')
male = sum(arms$Sex == 'Male')
female = sum(arms$Sex == 'Female')

male_folded = sum((arms$Sex == 'Male') & (arms$LonR_fold == 1))
female_folded = sum((arms$Sex == 'Female') & (arms$LonR_fold == 1))
male_folded_pct = round((male_folded / male), 2)
female_folded_pct = round((female_folded / female), 2)
obs_diff = male_folded_pct - female_folded_pct
cat("There are", male, "males and", female, "females.\n\n")
cat("Sample proportion of males who folded their left arm on top:", paste0(male_folded_pct), "\n\n")
cat("Sample proportion of females who folded their left arm on top:", paste0(female_folded_pct), "\n\n")
cat("Observed difference in proportions (males minus females):", paste0(obs_diff), "\n\n")
```

```{r Problem 1B, echo=FALSE, results = 'show'}
boot_arms = prop.test(LonR_fold~Sex, data= arms)
confint(boot_arms, level = 0.95)
# Hand calculated
p1 = round(male_folded/male, 4)
p2 = round(female_folded/female, 4)
SE_diff = sqrt((p1 * (1-p1)/male) + (p2 * (1 - p2)/female))
z_score = 1.96
obs_diff = p1 - p2
margin_of_error = z_score * SE_diff
lower_bound <- obs_diff - margin_of_error
upper_bound <- obs_diff + margin_of_error
```
Formula for Standard Error: $SE(\hat{p}_1 - \hat{p}_2) = \sqrt{\frac{\hat{p}_1(1 - \hat{p}_1)}{N_1} + \frac{\hat{p}_2(1 - \hat{p}_2)}{N_2}}$

Values plugged into the formula: $P_1$ = `r p1`, $P_2$ = `r p2`, $N_1$ = `r male`, $N_2$ = `r female`

Z-score used for 95% confidence: `r z_score`. This is because in a normal distribution, 95% of the data lies within around 1.96 standard deviations from the mean.

Standard error: `r round(SE_diff, 4)`

Margin of error: `r round(margin_of_error, 4)`

95% Confidence Interval for the difference in proportions: [`r round(lower_bound, 4)`, `r round(upper_bound, 4)`]

If we were to repeat this experiment many times, 95% of the time, the true difference in proportions (between males and females) would fall within the interval of [`r round(lower_bound, 4)`, `r round(upper_bound, 4)`].

The standard error represents how much the difference in proportions (between males and females) is likely to vary across different random samples drawn from the population. It measures the expected spread of the difference in proportions if we were to repeatedly sample from the population, as in bootstrapping.

A sampling distribution refers to distribution of the difference in sample proportions. It shows how the statistic would vary across samples. The proportion of males and females who folded their left arm over their right will vary, and therefore, so will the difference in the proportion of those two. This is due to the nature of random sampling. However, the true population proportions of the males and females who folded their left arm over their right stays fixed. This is the "true" values we would discover if we could somehow repeat this experiment for all men and women.

The mathematical theorem which justifies the use of a normal distribution is the Central Limit Theorem (heretofore referred to as the CLT). The CLT states that for sufficiently large sample sizes (NP >= 10 and N(1-P) >= 10), the sampling distribution of the sample mean/proportion is approximately normal, regardless of the shape of the population's distribution. Since $N_1$($P_1$) = `r male * p1`,  which is greater than or equal to 1010 and $N_1$(1-$P_1$) = `r male * (1-p1)`,  which is greater than or equal to 10, this theorem holds for the male proportions. Since $N_2$($P_2$) = `r female * p2`, which is greater than or equal to 10 and $N_2$(1-$P_2$) = `r female * (1-p2)`, which is greater than or equal to 10, this theorem holds for females too. Thus, it holds for the difference of the two sample proportions.

If the interval were [-0.01, 0.30], it would be fair to say that there is no  difference in arm folding between sexes. This is because the null hypothesis is there is no difference between the sexes, and due to 0 being inside this interval, we would fail to reject this hypothesis. Thus, there would not be statistical evidence of any difference.

If this experiment were conducted many times with different random samples of university students, it is highly likely that the confidence intervals would differ across samples. This is because each random sample introduces some variability in the sample difference in proportions, and as a result, the values will change with each new sample. In any random sample, there is a degree of randomness, and as such, the sample proportions will vary slightly. This variability is often referred to as Monte Carlo variation. However, we can be confident that approximately 95% of the confidence intervals would contain the true difference in proportions for the population (all men and women). 

```{r Problem 2, echo=FALSE, results = 'show', show_col_types = FALSE}
turnout = read_csv('turnout.csv')
called = turnout[turnout$GOTV_call == 1, ]
called_voted <- round(mean(called$voted1998 == 1), 4)
not_called = turnout[turnout$GOTV_call == 0, ]
not_called_voted = round(mean(not_called$voted1998 == 1), 4)
boot_turtnout = prop.test(voted1998 ~ GOTV_call, data= turnout)
confint(boot_turtnout, level = 0.95)
voted = turnout[turnout$voted1998 == 1,]
not_voted = turnout[turnout$voted1998 == 0,]
# Summarize data by age
called$GOTV_call <- NULL
not_called$GOTV_call <- NULL
called$PERSONS <- NULL
not_called_voted$PERSONS = NULL
voted$GOTV_call <- NULL
not_voted$GOTV_call <- NULL
voted$PERSONS <- NULL
not_voted$PERSONS = NULL
# Loop through each column in the 'called' dataset and calculate the mean
# Loop for calculating difference in means for each variable
con_age1 = mean(called$AGE, na.rm = TRUE)
con_majorpty1 = mean(called$MAJORPTY, na.rm = TRUE)
con_voted1996_1 = mean(called$voted1996, na.rm = TRUE)
con_age2 = mean(not_called$AGE, na.rm = TRUE)
con_majorpty2 = mean(not_called$MAJORPTY, na.rm = TRUE)
con_voted1996_2 = mean(not_called$voted1996, na.rm = TRUE)

# Mean of AGE, MAJORPTY, and voted1996 for those who voted
con_age_voted = mean(voted$AGE, na.rm = TRUE)
con_majorpty_voted = mean(voted$MAJORPTY, na.rm = TRUE)
con_voted1996_voted = mean(voted$voted1996, na.rm = TRUE)

# Mean of AGE, MAJORPTY, and voted1996 for those who did not vote
con_age_not_voted = mean(not_voted$AGE, na.rm = TRUE)
con_majorpty_not_voted = mean(not_voted$MAJORPTY, na.rm = TRUE)
con_voted1996_not_voted = mean(not_voted$voted1996, na.rm = TRUE)

```
The proportion of those receiving a GOTV call and that voted was: **`r called_voted`**

The proportion of those not receiving a GOTV call and that voted was: **`r not_called_voted`**

Those who received a call had a mean age of **`r round(con_age1, 4)`**, while those who did not had a mean age of **`r round(con_age2, 4)`**.

For those who voted, the mean age was **`r round(con_age_voted, 4)`**, and for those who did not vote, the mean age was **`r round(con_age_not_voted, 4)`**.

Those who received a call had a mean major political party membership proportion of **`r round(con_majorpty1, 4)`**, while those who did not had a mean major political party membership proportion of **`r round(con_majorpty2, 4)`**.

For those who voted, the mean major political party membership was **`r round(con_majorpty_voted, 4)`**, and for those who did not vote, the mean major political party membership was **`r round(con_majorpty_not_voted, 4)`**.

Those who received a call had a mean 1996 voting proportion of **`r round(con_voted1996_1, 4)`**, while those who did not had a mean 1996 voting proportion of **`r round(con_voted1996_2, 4)`**.

For those who voted, the mean 1996 voting proportion was **`r round(con_voted1996_voted, 4)`**, and for those who did not vote, the mean 1996 voting proportion was **`r round(con_voted1996_not_voted, 4)`**.

As evident by the presentation of summary statistics, both the likelihood of receiving a call and the voting outcome are influenced by these variables, indicating that they act as confounders in the relationship between the treatment (GOTV call) and the outcome (voting behavior).


```{r Problem 2b, echo=FALSE, results = 'hide', show_col_types = FALSE}
turnout_matched = matchit(GOTV_call ~ voted1996 + AGE + MAJORPTY,
                                 data = turnout,
                            method = 'optimal',
                           ratio = 5)
summary(turnout_matched)
matched_data <- match.data(turnout_matched)


matched_called = matched_data[matched_data$GOTV_call == 1, ]
matched_called_voted <- round(mean(matched_called$voted1998 == 1), 4)
matched_not_called = matched_data[matched_data$GOTV_call == 0, ]
matched_not_called_voted = round(mean(not_called$voted1998 == 1), 4)
matched_voted = matched_data[matched_data$voted1998 == 1,]
matched_not_voted = matched_data[matched_data$voted1998 == 0,]

matched_con_age1 = mean(matched_called$AGE, na.rm = TRUE)
matched_con_majorpty1 = mean(matched_called$MAJORPTY, na.rm = TRUE)
matched_con_voted1996_1 = mean(matched_called$voted1996, na.rm = TRUE)
matched_con_age2 = mean(matched_not_called$AGE, na.rm = TRUE)
matched_con_majorpty2 = mean(matched_not_called$MAJORPTY, na.rm = TRUE)
matched_con_voted1996_2 = mean(matched_not_called$voted1996, na.rm = TRUE)

# Mean of AGE, MAJORPTY, and voted1996 for those who voted
matched_con_age_voted = mean(matched_voted$AGE, na.rm = TRUE)
matched_con_majorpty_voted = mean(matched_voted$MAJORPTY, na.rm = TRUE)
matched_con_voted1996_voted = mean(matched_voted$voted1996, na.rm = TRUE)

# Mean of AGE, MAJORPTY, and voted1996 for those who did not vote
matched_con_age_not_voted = mean(matched_not_voted$AGE, na.rm = TRUE)
matched_con_majorpty_not_voted = mean(matched_not_voted$MAJORPTY, na.rm = TRUE)
matched_con_voted1996_not_voted = mean(matched_not_voted$voted1996, na.rm = TRUE)


```


The proportion of those receiving a GOTV call and that voted was: **`r round(matched_called_voted, 4)`**

The proportion of those not receiving a GOTV call and that voted was: **`r round(matched_not_called_voted, 4)`**

Those who received a call had a mean age of **`r round(matched_con_age1, 4)`**, while those who did not had a mean age of **`r round(matched_con_age2, 4)`**.

Those who received a call had a mean major political party membership proportion of **`r round(matched_con_majorpty1, 4)`**, while those who did not had a mean major political party membership proportion of **`r round(matched_con_majorpty2, 4)`**.

Those who received a call had a mean 1996 voting proportion of **`r round(matched_con_voted1996_1, 4)`**, while those who did not had a mean 1996 voting proportion of **`r round(matched_con_voted1996_2, 4)`**.

Since these variables still influence the outcome, they remain relevant predictors. However, because the treatment assignment is no longer systematically affected by them due to matching, they are no longer confounding variables.

```{r Problem 2c, echo=FALSE, results = 'show', show_col_types = FALSE}
boot_matched_turtnout = prop.test(voted1998 ~ GOTV_call, data= matched_data)
confint(boot_matched_turtnout, level = 0.95)
```
Because the 95% confidence interval does not include 0, this allows us to reject the null hypothesis (that there was no difference in voter turnout between those who received the call and those who did not), **therefore, receiving a GOTV call did have an effect on voter turnout, increasing the likelihood that a person voted.**



```



